---
title: "PGLS - ADH Loss & Diet"
date: "1/24/2020"
output: html_document
---

Load libraries:
```{r}
library(ape)
library(geiger)
library(nlme)
library(phytools)
library(phylolm)
library(ggplot2)
library(reshape2)
library(wesanderson)
```


Read in data and tree
```{r}
adh_tree <- read.tree("Mammal_ADH_tree.nwk")
adh_data <- read.csv("ADH_Diet_PGLS.tsv", header = TRUE, sep = "\t", row.names = 1)
```

Check if tree names and data names match:
```{r}
name.check(adh_tree, adh_data)
```


## phyloglm analyses:

Prep data by creating binary column for gene status (pseudo = 1, functional = 0)
```{r}
adh_data$ADH7_binary <- 0
adh_data$ADH7_binary[adh_data$ADH7 == 'pseudo'] <- 1
```

### Testing hypothesis that herbivory predicts gene loss:

* Method logistic_MPLE
```{r}
plant_MPLE <- phyloglm(ADH7_binary ~ perc_plant, data = adh_data, phy = adh_tree, method = "logistic_MPLE", btol = 20)
summary(plant_MPLE)
coef(plant_MPLE)
vcov(plant_MPLE)
```

* Method logistic_IG10
```{r}
plant_IG10 <- phyloglm(ADH7_binary ~ perc_plant, data = adh_data, phy = adh_tree, method = "logistic_IG10", btol = 10)
summary(plant_IG10)
# coef(plant_IG10)
vcov(plant_IG10)
```

predict lines for plot:
```{r}
cc_plant_MPLE <- coef(plant_MPLE)
cc_plant_IG10 <- coef(plant_IG10)
```

plot:
```{r}
plot(adh_data$ADH7_binary ~ adh_data$perc_plant)
curve_plant_MPLE <- curve(plogis(cc_plant_MPLE[1]+cc_plant_MPLE[2]*x),col="red",add=TRUE)
curve_plant_IG10 <- curve(plogis(cc_plant_IG10[1]+cc_plant_IG10[2]*x),col="red",add=TRUE)

```

In ggplot:
Prep data
```{r}
df_plant_curve <- data.frame(X=curve_plant_MPLE$x, MPLE=curve_plant_MPLE$y, IG10=curve_plant_IG10$y)
df_plant_curve_melted <- melt(df_plant_curve, id.vars = "X")
names(df_plant_curve_melted)[2] <- "Method"
```
Plot:
```{r}
plot <- ggplot(data = adh_data, aes(x=perc_plant, y = ADH7_binary)) + geom_count() + theme_classic() + xlab("Vegetative material in diet (%)") + ylab("ADH7") +
scale_y_continuous(breaks = c(0.00,1.00), labels = c("Retention","Loss")) + theme(axis.title.y = element_text(face = 'italic')) + labs(size = "Number of species") + scale_size_continuous(breaks = c(1,5,9,13))
plot + geom_line(data = df_plant_curve_melted, aes(x=X, y=value, col=Method)) + scale_color_manual(values = c(wes_palette("Zissou1")[3], wes_palette("Zissou1")[1])) + geom_label(label = "slope = 0.045\np < 0.004", x=75, y=0.73, size = 3, fill = wes_palette("Zissou1")[1]) + geom_label(label = "slope = 0.042\np = 0.005", x=85, y=0.44, size = 3, fill = wes_palette("Zissou1")[3])
#add labels for coefficients of the two models
```

### Testing hypothesis that frugivory predicts gene retention:
Prep data by creating binary column (pseudo = 1, functional = 0)
```{r}
adh_data$ADH7_binary <- 0
adh_data$ADH7_binary[adh_data$ADH7 == 'pseudo'] <- 1
```

```{r}
fruit_MPLE <- phyloglm(ADH7_binary ~ perc_fruit_nectar, data = adh_data, phy = adh_tree, method = "logistic_MPLE", btol = 20)
summary(fruit_MPLE)
coef(fruit_MPLE)
vcov(fruit_MPLE)
```

```{r}
fruit_IG10 <- phyloglm(ADH7_binary ~ perc_fruit_nectar, data = adh_data, phy = adh_tree, method = "logistic_IG10", btol = 10)
summary(fruit_IG10)
coef(fruit_IG10)
vcov(fruit_IG10)
```

Predict lines for plot:
```{r}
cc_fruit_MPLE <- coef(fruit_MPLE)
cc_fruit_IG10 <- coef(fruit_IG10)
```
plot:
```{r}
plot(adh_data$ADH7_binary ~ adh_data$perc_fruit_nectar)
curve_fruit_MPLE <- curve(plogis(cc_fruit_MPLE[1]+cc_fruit_MPLE[2]*x),col="red",add=TRUE)
curve_fruit_IG10 <- curve(plogis(cc_fruit_IG10[1]+cc_fruit_IG10[2]*x),col="red",add=TRUE)
```

In ggplot:
Prep data
```{r}
df_fruit_curve <- data.frame(X=curve_fruit_MPLE$x, MPLE=curve_fruit_MPLE$y, IG10=curve_fruit_IG10$y)
df_fruit_curve_melted <- melt(df_fruit_curve, id.vars = "X")
names(df_fruit_curve_melted)[2] <- "Method"
```
Plot:
```{r}
plot <- ggplot(data = adh_data, aes(x=perc_fruit_nectar, y = ADH7_binary)) + geom_count() + theme_classic() + xlab("Fruit and/or nectar in diet (%)") + ylab("ADH7") +
scale_y_continuous(breaks = c(0.00,1.00), labels = c("Retention","Loss")) + theme(axis.title.y = element_text(face = 'italic')) + labs(size = "Number of species") + scale_size_continuous(breaks = c(2,4,6,9))
plot + geom_line(data = df_fruit_curve_melted, aes(x=X, y=value, col=Method)) + scale_color_manual(values = c(wes_palette("Zissou1")[3], wes_palette("Zissou1")[1])) + geom_label(label = "slope = 0.000\np = 1.000", x=75, y=0.73, size = 3, fill = wes_palette("Zissou1")[1]) + geom_label(label = "slope = 0.000\np = 1.000", x=85, y=0.44, size = 3, fill = wes_palette("Zissou1")[3])
```


PGLS analyses: (not included in paper)
* Are number of missense mutations and being an herbivore correlated?
```{r}
pglsModel <- gls(mutations_stop ~ herbivore, correlation = corBrownian(phy = adh_tree), data = adh_data, method = "ML")
summary(pglsModel)
coef(pglsModel)
anova(pglsModel)
```

Are number of missense mutations and percent fruit/nectar in diet correlated? (not included in paper)
```{r}
fruit_pglsModel <- gls(mutations_stop ~ perc_fruit_nectar, correlation = corBrownian(phy = adh_tree), data = adh_data, method = "ML")
summary(fruit_pglsModel)
coef(fruit_pglsModel)
anova(fruit_pglsModel)
```


